---
title: "HARplus: Enhanced Tools for GEMPACK File Processing"
author: "Pattawee Puangchit"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{HARplus: Enhanced Tools for GEMPACK File Processing}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE
)
```

# Introduction

This vignette demonstrates the use of the `group_by_dims` and `group_by_dims_from_sl4` functions, which facilitate structured data grouping for GEMPACK users. The package was initially developed based on the GTAPv7 model and utilizes the freely available GTAP 9 database.  
For more details on the GTAP Data Base, refer to the official archive:  
[https://www.gtap.agecon.purdue.edu/databases/archives.asp](https://www.gtap.agecon.purdue.edu/databases/archives.asp).

The examples in this vignette utilize two **simulation solution files** (`.sl4`) and two **welfare decomposition files** (`.har`). The primary focus is to showcase how results can be processed using this package, without interpreting simulation outcomes or discussing model setup.

First, you can download and install the package then calling the example files using command:

```{r load-package}
devtools::install("D:/GitHub/HARplus")
library(HARplus)

har_file <- system.file("data", "ExB14-WEL.har", package = "HARplus")
sl4_file <- system.file("data", "ExB14.sl4", package = "HARplus")
```
